Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Istanbul version \YellowPaperVersionNumber}}

### 2021-06-14 Paper.tex
EIP-3607: transactions from senders with deployed contract are forbidden
Creation on non-empty code or non-zero nonce throws an exception (#323)
Co-authored-by: Andrew Ashikhmin <34320705+yperbasis@users.noreply.github.com>
Amend contract creation when initialising code reverts

\begin{align}
\quad g' &\equiv \begin{cases}
0 & \text{if} \quad F \\
g^{**} - c & \text{otherwise} \\
\end{cases} \\
\quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
\boldsymbol{\sigma} & \text{if} \quad F \\
\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}^{**}, a) \\
\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{KEC}(\mathbf{o}) & \text{otherwise}
\end{cases} \\
\quad z &\equiv \begin{cases}
0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
1 & \text{otherwise}
\end{cases} \\
\nonumber \text{where} \\
F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  \lVert \mathbf{o} \rVert > 24576\big)
\end{align}

@@

Note that intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.

@@

Aside from evaluating to a new state and accrued transaction substate, message calls also have an extra component---the output data denoted by the byte array $\mathbf{o}$. This is ignored when executing transactions, however message calls can be initiated due to VM-code execution and in this case this information is used.
\begin{equation}
(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \linkdest{theta}{\Theta}(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
\end{equation}
Note that we need to differentiate between the value that is to be transferred, $v$, from the value apparent in the execution context, $\tilde{v}$, for the {\small DELEGATECALL} instruction.

@@

0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\

>>>>

\begin{align}
\quad g' \equiv & \begin{cases}
0 & \text{if} \quad F \\
g^{**} - c & \text{otherwise} \\
\end{cases} \\
\quad \boldsymbol{\sigma}' \equiv & \begin{cases}
\boldsymbol{\sigma} & \text{if} \quad F \ \lor\ \boldsymbol{\sigma}^{**} = \varnothing \\
\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}^{**}, a) \\
\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{KEC}(\mathbf{o}) & \text{otherwise}
\end{cases} \\
\quad z \equiv & \begin{cases}
0 & \text{if} \quad F \ \lor\ \boldsymbol{\sigma}^{**} = \varnothing \\
1 & \text{otherwise}
\end{cases} \\
\nonumber \text{where} \\
F \equiv  & \big( \boldsymbol{\sigma}[a] \neq \varnothing \ \wedge\ \big(\boldsymbol{\sigma}[a]_c \neq \texttt{\small KEC}\big(()\big) \vee \boldsymbol{\sigma}[a]_n \neq 0 \big) \big) \quad \vee \\
\nonumber &(\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \quad \vee \\
\nonumber &g^{**} < c \quad \vee \\
\nonumber &\lVert \mathbf{o} \rVert > 24576
\end{align}

@@

Note that intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
In addition, observe that if the execution of the initialising code \hyperlink{REVERT}{reverts} ($\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} \neq \varnothing$), the resultant gas $g'$ is not depleted (provided there was no other exception), but no new account is created.

@@

Aside from evaluating to a new state and accrued transaction substate, message calls also have an extra component---the output data denoted by the byte array~$\mathbf{o}$. This is ignored when executing transactions, however message calls can be initiated due to VM-code execution and in this case this information is used.
\begin{equation}
(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \linkdest{theta}{\Theta}(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
\end{equation}
Note that we need to differentiate between the value that is to be transferred, $v$, from the value apparent in the execution context, $\tilde{v}$, for the {\small DELEGATECALL} instruction.

@@

\linkdest{REVERT}{}0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
